The Analog Method for Precipitation Forecasting: Finding Better Analog Situations at a Sub-Daily Time Step

### 2016-08-17 Moving temporal window - HESS/template.tex
Changing "analogue" for "analog".

\title{The Analogue Method for Precipitation Forecasting: Finding Better Analogue Situations at a Sub-Daily Time Step}

@@

\runningtitle{Finding Better Analogue Situations at a Sub-Daily Time Step}

@@

\begin{abstract}
The Analogue Method (AM) aims at forecasting local weather variables (predictands), such as precipitations, by means of a statistical relationship with predictors at a synoptic scale. The analogy is generally assessed in the first place on the geopotential field by mean of a comparison of the gradients, in order to sample the days with a similar atmospheric circulation.

The search for candidate situations, for a given target day, is usually undertaken by comparing the state of the atmosphere at fixed hours of the day, for both the target day and the candidate analogues. The constraint being the use of daily time series, due to the length of available archives they provide, and the unavailability of equivalent archives at a finer time step. However, it is unlikely that the best analogy happens at the very same hour, but it may occur at a different time of the day. In order to assess the potential of finding better analogues at a different hour, a moving time window (MTW) has been introduced on a reduced archive of hourly precipitation totals.

The MTW resulted in a better analogy in terms of the atmospheric circulation, with improved values of the analogy criteria on the whole distribution of analogue dates. The improvement was found to grow with the analogue ranks due to an accumulation of more similar situations in the selection. Moreover, the improvement is even more important for days with heavy precipitation events, which are generally related to more dynamic atmospheric situations, where timing is more specific.

A seasonal effect has also been identified, with larger improvements in winter than in summer, supposedly due to the stronger effect of the diurnal cycle in summer, which favors predictors at the same hour for target and analogues.

@@

The Analogue Method (AM) is based on the principle that two relatively similar synoptic situations may produce relatively similar local effects \citep{Lorenz1956, Lorenz1969}. Multiple variations of the methods exist and listings can be found in \citet{Horton2016} or \citet{BenDaoud2015}. The versions that will be considered here are often used as references or benchmarks for various improvements.

The first method is based on the analogy of the atmospheric circulation only \citep[Table \ref{table:method_2Z},][]{Obled2002, Bontron2005, Marty2012}. Searching analogue situations for a target day starts by a preselection step of the potential candidate for analogy. Here, it has been limited to the 4 months centered around the target date for every year of the archive, in order to cope with seasonal effects. Then, the similarity of the atmospheric circulation of the target date with every day of the preselection is assessed by processing the \citet{Teweles1954} score ($S_{\text{TW}}$), which is a comparison of gradients, over a certain spatial window and at certain hours. The smaller the $S_{\text{TW}}$ values, the more similar the pressure fields.

@@

Then, $N_{1}$ dates with the lowest values of $S_{\text{TW}}$ are considered as analogues to the target day, $N_{1}$ being a parameter to calibrate. Then, the daily observed precipitation amount of the corresponding dates provide the empirical conditional distribution considered as the probabilistic forecast for the target day. Following the nomenclature proposed by \citet{Horton2016}, this method will be named 2Z.

The second reference method adds a subsequent level of analogy on moisture variables, compared by means of the root-mean-square error ($E_{\text{RMS}}$) (method 2Z-2MI, Table \ref{table:method_2Z-2MI}). This second predictor is a moisture index made of the product of the total precipitable water (TPW) with the relative humidity at 850~hPa (RH850) \citep{Bontron2004}. When adding a second level of analogy, $N_{2}$ dates are subsampled in the $N_{1}$ analogues on the atmospheric circulation, to end up with a smaller number of analogue situations. When a second level of analogy is added, a higher number of $N_{1}$ analogues is kept on the first level.

To calibrate the method, the commonly used technique is a semi-automatic sequential procedure elaborated by \citet{Bontron2004}. AtmoSwing \citep{Horton2016}, which implements this technique, was used in order to calibrate the parameters in the present study. The score that is most often considered to assess an AM performance is the Continuous Ranked Probability Score \citep[$S_{\text{CRP}}$,][]{Brown1974, Matheson1976, Hersbach2000}. It allows evaluating the predicted cumulative distribution functions $F(y)$, for example of the precipitation values $y$ from analogue situations, compared to the observed value $y^{0}$. The better the forecast, the smaller the score. It is defined as follows:

@@

The AM is usually implemented with a daily time step, due to the availability of long precipitation archives that have no equivalent at a finer resolution. Therefore, the analogue situations are assessed on the basis of a daily time step, by comparing predictors at fixed hours of the day, otherwise one would not know what precipitation values to assign to them. However, it can be expected that the analogy of the synoptic situations does not occur systematically at the same time of the day, and that better candidates can be found by shifting to a different hour. On this assumption, a moving time window (MTW) was introduced to allow searching for candidates at different hours of the day.

Previous tests showed the benefit, in terms of analogy criteria values, of searching for analogue synoptic situations at a finer time step, but without assessing the impact on the forecast skills \citep{Finet2008}. In order to assess this potential improvement of the prediction, precipitation data with a resolution of 10 minutes, on a respectable archive length, was used (Sect. \ref{sec:data}). It was then aggregated in the form of 24~h-totals, but starting at different 6-hourly time step, by means of a moving total (Sect. \ref{sec:method}).

@@

The study area is the upper Rh\^{o}ne catchment in Switzerland \cite[see also][]{Horton2012a}. Due to the low density of weather stations with high temporal resolution and long archives, no spatially aggregated rainfall was processed. The time series (on the period 1982-2007) come from 6 automatic weather stations, namely Ulrichen, Zermatt, Visp, Montana, Sion and Aigle (Fig. \ref{fig:map}), which are subject to various meteorological influences \citep{Horton2012}. The results will thereafter be presented arbitrarily for the Ulrichen station, but they are equivalent for all stations. The precipitation time series were then aggregated by means of a moving 24~h-total (see next section).

@@

In order to assess the benefit of searching analogue situations at a sub-daily time step, an appropriate precipitation series is required. On the basis of high resolution time series (Sect. \ref{sec:data}), 24~h-totals were processed, but at a 6-hourly time step (temporal resolution of the reanalysis dataset), by means of a moving 24~h-total.

The target situations and their corresponding observed precipitation values (used for validation) do not change, because the forecast is always established for a fixed period of the target day (6-30~h), as before. The difference is that the candidates are 4~times more numerous (Fig. \ref{fig:principle}).

Based on some preliminary tests (not shown), no constraint was added in order to restrict the selection of multiple analogues within the same candidate dates.

@@

When searching for analogues on the geopotential heights, as in the 2Z method, there are now 4 times more candidates than before, which obviously allows to find better matches.

Figure \ref{fig:changes_S1_analogues} presents the changes in the distributions of the $S_{\text{TW}}$ criterion for the $1^{st}$, $5^{th}$, $20^{th}$ and $40^{th}$ analogues for the Ulrichen station on the whole calibration period, due to introduction of the MTW. The precipitation target remains as before, that is centered on 18~h~UTC (6~h~UTC to 6~h~UTC the next day). The shapes of the distributions of the conventional approach and the MTW are similar, but the values of the analogy criteria are now reduced (shifted to the left), and therefore better. An increase in the difference between a fixed window and a moving window is identifiable, which means that the last analogues are further improved. The latter effect is due to the accumulation of improvements brought by the new analogue situations in the selection.

The improvements of the $S_{\text{TW}}$ criteria are summarized in Fig. \ref{fig:changes_S1}, which shows (top) quantiles of the $S_{\text{TW}}$ criteria according to the analogue rank for the conventional method and the MTW, and (bottom) quantiles of the relative reduction. This confirms that all quantile seem similarly reduced ($S_{\text{TW}}$ distributions keep their shape), and that this improvement is constantly increasing from the first to the last analogue (Fig. \ref{fig:changes_S1} bottom).

The median of the $S_{\text{TW}}$ values reduction (Fig. \ref{fig:changes_S1} bottom) starts approximately at 5~\% for the first analogue and reaches more than 10~\% for the last one. This increasing trend with the analogue rank can be explained by the accumulation of better analogues in the distribution. The minimum improvement starts from 0 and reaches 2-3\%, meaning that the criteria have been improved on most analogues for every day of the calibration period. All other stations have a similar improvement of the $S_{\text{TW}}$ criteria, both in terms of distribution shape and amplitude.

@@

A seasonal effect can be observed on the reduction of the $S_{\text{TW}}$ criteria due to the MTW (Fig. \ref{fig:changes_S1_seasons}). The improvements are greater for winter than summer. One hypothesis is that the diurnal effects of the summer months have an influence on the atmospheric circulation, at least in the lower layers. This effect is based on the daily cycle and good analogues are essentially found for the same hours or time windows.

An analysis of the selected hours for the geopotential predictor seems to confirm this assumption (Fig. \ref{fig:hours_selection_per_season}). It was found that the new choice of the temporal window in winter, when using the MTW approach, is well balanced between the 4 options. This means a change of 75\% of the analogues selection compared to the conventional approach, which improves the circulation analogy.

On the contrary, the summer months have a preference for the initial temporal window (Z500 24h \& Z1000 12h), due to more pronounced diurnal effects which reduces the potential for improvement of the criteria. Other seasons are between these two extremes, which is consistent with their respective improvements. This seasonal effect was observed for each station in a very similar way, and generally even with a slightly larger amplitude than for Ulrichen (see example of the Visp station in Fig. \ref{fig:hours_selection_per_season_Visp}).

@@

When considering the second level of analogy of the 2Z-2MI method (Table \ref{table:method_2Z-2MI}), the number of candidates situations did not increase, as they are subsampled in the previously $N_{1}$ selected analogues, but their dates have changed. In contrast to earlier, both a reduction or an increase of the $E_{\text{RMS}}$ analogy criterion values are possible. Indeed, Fig. \ref{fig:changes_RMSE} shows an almost insignificant improvement of the $E_{\text{RMS}}$ values. Unlike for the first level of analogy, the relative improvement of the $E_{\text{RMS}}$ values are distributed relatively symmetrically around zero, with improvements and losses of the same amplitude. Once again, the results for the other stations are similar.

This result of a globally null improvement of the $E_{\text{RMS}}$ values does not mean that the 2Z-2MI method cannot be improved by the MTW. It means that after the selection of the analogues situations in terms of the synoptic circulation, the new candidate dates do not allow to find better analogues in terms of moisture. However, the selected dates have changed in the first level of analogy, and thus also in the final selection, which can potentially improve the performance scores on the predictand.

@@

A systematic improvement of the $S_{\text{TW}}$ values was previously observed. However, finding better analogue situations does not obligatory imply better skills to predict the precipitation. Therefore, the impact of the MTW introduction, and thus the selection of other analogue dates, has to be assessed on the performance scores. In order to perform this assessment, the 24~h-totals (moving average) at a 6-hourly time step were used (see Sect. \ref{sec:data}). The target dates remain unchanged, since the original time slot (6~h~UTC - 6~h~UTC the next day) is kept, and thus the performance scores can be directly compared with the former ones.

@@

After recalibrating, some changes in the optimal parameters can be observed for both 2Z (Table \ref{table:params_2Z_new}) and 2Z-2MI methods (Table \ref{table:params_2Z-2MI_new}). Among these, the east-west dimension of the spatial windows of the circulation analogy tends to decrease. More significantly, the optimal number of analogues increases after introducing the MTW, of a significant order of magnitude: 25~\% to 83~\% for the 2Z method and 20~\% 67~\% for the 2Z-2MI method. The number of analogues of the first analogy level of the 2Z-2MI method even reached three times its previous value for the Visp station. It seems as if the method perceive the increase in the archive length and diversity by allowing to extract more analogues. As we saw in Fig. \ref{fig:changes_S1}, the improvement of the $S_{\text{TW}}$ criteria grows along with the rank of the analogue, which shows an accumulation of better analogue situations in the distributions. It seems thus profitable to widen the selection of analogues in order to keep also some whose rank has increased, as they appear to be also relevant to predict precipitation values. The number of good analogues is thus globally increased.

This increase in number of analogues has a slight effect on the performance of the different precipitation classes. The same analysis as in Sect. \ref{sec:improvement_CRPSS_precip_threshold} has been performed again on the basis of the newly calibrated parameters. The results are generally very similar, but a slight performance increase of small precipitation values can be observed at the expense of higher amounts. The change in the analogues numbers is likely to be responsible for this difference in the performance distributions.

@@

A proportional distribution is certainly the simplest approach that can be performed. It consists in allocating proportional parts of the original daily time series into a new moving average of 24~h-totals (Fig. \ref{fig:illustration_disaggregation}). When using this reconstructed time series on the shorter period, the method performance was degraded (Table \ref{table:disaggregation_proportional} to compare to Table \ref{table:CRPSS_MTW}) and was even below the conventional method without MTW (Table \ref{table:loss_reduction}). The benefit of a better selection of the analogue situations is lost due to a precipitation archive of poor quality.

@@

The first step consisted in determining which variables from the precipitable water or the relative humidity (at 1000~hPa, 925~hPa or 850~hPa) produces the precipitation chronology that is the most correlated (on non-zero values) with the accurate 6-hourly precipitation time series on the period 1982-2007. The location of the most predictive point from the reanalysis grid had to be assessed (closest points: 5\textdegree\ and 7.5\textdegree\ longitude and 45\textdegree\ and 47.5\textdegree\ latitude), as well an eventual time lapse between both series, due to the significant distance separating the weather stations and the reanalysis grid point. The most relevant point will probably not be the same for all stations, since one can expect to find the optimum in the direction of the main moisture inflows for the station. This assessment must be carried out for each station.

The results of the correlation analyses (on non-zero values) between the new series reconstituted using the proxy and the actual series are illustrated in Table \ref{table:proxy_correlations} for Zermatt (on the short period, 1982-2007). The best proxy is the precipitable water at 45\textdegree\ latitude and 7.5\textdegree\ longitude, with a time offset of -6~h. The determination coefficient of 0.721 is higher than that of the series from the proportional distribution (0.698), which confirms that a little information was added to the series of precipitation.

@@

The AM is most often based on a daily time step, due to the availability of long precipitation archives. However, it is unlikely that two analogue synoptic situations, that evolve relatively quickly, would correspond optimally at the same hour of the day. It is thus more likely that a better match can be found at another time, which can change the selection of the analogues dates.

As \citet{Finet2008} had previously shown, the introduction of a MTW allows finding better analogue situations in terms of the atmospheric circulation. It has been here demonstrated that the improvement of the $S_{\text{TW}}$ criteria values is growing with the rank of the analogue. This is due to the accumulation of better analogues within the predicted distributions.

@@

A seasonal effect has been highlighted, as the MTW is more profitable for winter months than summer. The reason is likely that the diurnal cycle has a bigger effect in summer than in winter, which results in better analogues at the same time of the day. The preference for the same hours in summer has been demonstrated, whereas 75\% of the analogue situations are selected at a different time in winter.

@@

The parameters were then calibrated again, using the MTW. Some parameters changed, the main difference being the number of analogues, which systematically and significantly increased compared to the original set. A wider selection of analogue situations, containing those whose rank decreased, seemed profitable for the prediction performance. The number of good analogues is thus globally increased. This change seems to benefit mainly the prediction of days with small precipitation totals.

The importance of the quality of the precipitation archive was also demonstrated, as too simplistic reconstructions of 6-hourly time series lead to a loss of all the improvement brought by better analogue situations identified by the MTW. The precipitation prediction is improved only when the precipitation chronology is close to the accurate one. Attempts to reconstruct longer time series based on a simplistic proportional distribution or using meteorological variables from the NCEP reanalysis 1 dataset as proxy did not succeed. Other reanalysis datasets could eventually preform better.

@@

\begin{figure*}[htb]
	\begin{center}
		\includegraphics[width=15cm]{figures/changes_S1_analogues.pdf}
	\end{center}
	\caption{Changes in the $S_{\text{TW}}$ criteria distributions of the $1^{st}$, $5^{th}$, $20^{th}$ and $40^{th}$ analogue ranks for the Ulrichen station, due to the MTW.}
	\label{fig:changes_S1_analogues}
\end{figure*}

\begin{figure}[htb]
	\begin{center}
		\includegraphics[width=8.2cm]{figures/changes_S1_value.pdf} \\
		\includegraphics[width=8.3cm]{figures/changes_S1_gain.pdf}
	\end{center}
	\caption{Synthesis of the changes in the $S_{\text{TW}}$ criteria, due to the MTW, for the Ulrichen station, depending on the ranks of the analogue. (top) Quantiles of the $S_{\text{TW}}$ distributions with and without the MTW. (bottom) Quantiles of the relative improvements of the S1 criteria when using the MTW.}
	\label{fig:changes_S1}
\end{figure}

@@

\begin{figure}[htb]
	\includegraphics[width=8.3cm]{figures/hours_selection_per_season.pdf}
	\caption{Distribution of the predictors hours in the selected analogue dates depending on the season, for the Ulrichen station.}
	\label{fig:hours_selection_per_season}
\end{figure}

\begin{figure}[htb]
	\includegraphics[width=8.3cm]{figures/hours_selection_per_season_Visp.pdf}
	\caption{Distribution of the predictors hours in the selected analogue dates depending on the season, for the Visp station.}
	\label{fig:hours_selection_per_season_Visp}
\end{figure}

\begin{figure}[htb]
	\begin{center}
		\includegraphics[width=8.1cm]{figures/changes_RMSE_value.pdf} \\
		\includegraphics[width=8.3cm]{figures/changes_RMSE_gain.pdf}
	\end{center}
	\caption{Synthesis of the changes in the $E_{\text{RMS}}$ criteria, due to the MTW, for the Ulrichen station, depending on the ranks of the analogue. (top) Quantiles of the $E_{\text{RMS}}$ distributions with and without the MTW. (bottom) Quantiles of the relative improvements of the $E_{\text{RMS}}$ criteria when using the MTW.}
	\label{fig:changes_RMSE}
\end{figure}

>>>>

\title{The Analog Method for Precipitation Forecasting: Finding Better Analog Situations at a Sub-Daily Time Step}

@@

\runningtitle{Finding Better Analog Situations at a Sub-Daily Time Step}

@@

\begin{abstract}
The Analog Method (AM) aims at forecasting local weather variables (predictands), such as precipitations, by means of a statistical relationship with predictors at a synoptic scale. The analogy is generally assessed in the first place on the geopotential field by mean of a comparison of the gradients, in order to sample the days with a similar atmospheric circulation.

The search for candidate situations, for a given target day, is usually undertaken by comparing the state of the atmosphere at fixed hours of the day, for both the target day and the candidate analogs. The constraint being the use of daily time series, due to the length of available archives they provide, and the unavailability of equivalent archives at a finer time step. However, it is unlikely that the best analogy happens at the very same hour, but it may occur at a different time of the day. In order to assess the potential of finding better analogs at a different hour, a moving time window (MTW) has been introduced on a reduced archive of hourly precipitation totals.

The MTW resulted in a better analogy in terms of the atmospheric circulation, with improved values of the analogy criteria on the whole distribution of analog dates. The improvement was found to grow with the analog ranks due to an accumulation of more similar situations in the selection. Moreover, the improvement is even more important for days with heavy precipitation events, which are generally related to more dynamic atmospheric situations, where timing is more specific.

A seasonal effect has also been identified, with larger improvements in winter than in summer, supposedly due to the stronger effect of the diurnal cycle in summer, which favors predictors at the same hour for target and analogs.

@@

The Analog Method (AM) is based on the principle that two relatively similar synoptic situations may produce relatively similar local effects \citep{Lorenz1956, Lorenz1969}. Multiple variations of the methods exist and listings can be found in \citet{Horton2016} or \citet{BenDaoud2015}. The versions that will be considered here are often used as references or benchmarks for various improvements.

The first method is based on the analogy of the atmospheric circulation only \citep[Table \ref{table:method_2Z},][]{Obled2002, Bontron2005, Marty2012}. Searching analog situations for a target day starts by a preselection step of the potential candidate for analogy. Here, it has been limited to the 4 months centered around the target date for every year of the archive, in order to cope with seasonal effects. Then, the similarity of the atmospheric circulation of the target date with every day of the preselection is assessed by processing the \citet{Teweles1954} score ($S_{\text{TW}}$), which is a comparison of gradients, over a certain spatial window and at certain hours. The smaller the $S_{\text{TW}}$ values, the more similar the pressure fields.

@@

Then, $N_{1}$ dates with the lowest values of $S_{\text{TW}}$ are considered as analogs to the target day, $N_{1}$ being a parameter to calibrate. Then, the daily observed precipitation amount of the corresponding dates provide the empirical conditional distribution considered as the probabilistic forecast for the target day. Following the nomenclature proposed by \citet{Horton2016}, this method will be named 2Z.

The second reference method adds a subsequent level of analogy on moisture variables, compared by means of the root-mean-square error ($E_{\text{RMS}}$) (method 2Z-2MI, Table \ref{table:method_2Z-2MI}). This second predictor is a moisture index made of the product of the total precipitable water (TPW) with the relative humidity at 850~hPa (RH850) \citep{Bontron2004}. When adding a second level of analogy, $N_{2}$ dates are subsampled in the $N_{1}$ analogs on the atmospheric circulation, to end up with a smaller number of analog situations. When a second level of analogy is added, a higher number of $N_{1}$ analogs is kept on the first level.

To calibrate the method, the commonly used technique is a semi-automatic sequential procedure elaborated by \citet{Bontron2004}. AtmoSwing \citep{Horton2016}, which implements this technique, was used in order to calibrate the parameters in the present study. The score that is most often considered to assess an AM performance is the Continuous Ranked Probability Score \citep[$S_{\text{CRP}}$,][]{Brown1974, Matheson1976, Hersbach2000}. It allows evaluating the predicted cumulative distribution functions $F(y)$, for example of the precipitation values $y$ from analog situations, compared to the observed value $y^{0}$. The better the forecast, the smaller the score. It is defined as follows:

@@

The AM is usually implemented with a daily time step, due to the availability of long precipitation archives that have no equivalent at a finer resolution. Therefore, the analog situations are assessed on the basis of a daily time step, by comparing predictors at fixed hours of the day, otherwise one would not know what precipitation values to assign to them. However, it can be expected that the analogy of the synoptic situations does not occur systematically at the same time of the day, and that better candidates can be found by shifting to a different hour. On this assumption, a moving time window (MTW) was introduced to allow searching for candidates at different hours of the day.

Previous tests showed the benefit, in terms of analogy criteria values, of searching for analog synoptic situations at a finer time step, but without assessing the impact on the forecast skills \citep{Finet2008}. In order to assess this potential improvement of the prediction, precipitation data with a resolution of 10 minutes, on a respectable archive length, was used (Sect. \ref{sec:data}). It was then aggregated in the form of 24~h-totals, but starting at different 6-hourly time step, by means of a moving total (Sect. \ref{sec:method}).

@@

The study area is the upper Rh\^{o}ne catchment in Switzerland \cite[see also][]{Horton2012a}. Due to the low density of weather stations with high temporal resolution and long archives, no spatially aggregated rainfall was processed. The time series (on the period 1982--2007) come from 6 automatic weather stations, namely Ulrichen, Zermatt, Visp, Montana, Sion and Aigle (Fig. \ref{fig:map}), which are subject to various meteorological influences \citep{Horton2012}. The results will thereafter be presented arbitrarily for the Ulrichen station, but they are equivalent for all stations. The precipitation time series were then aggregated by means of a moving 24~h-total (see next section).

@@

In order to assess the benefit of searching analog situations at a sub-daily time step, an appropriate precipitation series is required. On the basis of high resolution time series (Sect. \ref{sec:data}), 24~h-totals were processed, but at a 6-hourly time step (temporal resolution of the reanalysis dataset), by means of a moving 24~h-total.

The target situations and their corresponding observed precipitation values (used for validation) do not change, because the forecast is always established for a fixed period of the target day (6--30~h), as before. The difference is that the candidates are 4~times more numerous (Fig. \ref{fig:principle}).

Based on some preliminary tests (not shown), no constraint was added in order to restrict the selection of multiple analogs within the same candidate dates.

@@

When searching for analogs on the geopotential heights, as in the 2Z method, there are now 4 times more candidates than before, which obviously allows to find better matches.

Figure \ref{fig:changes_S1_analogs} presents the changes in the distributions of the $S_{\text{TW}}$ criterion for the $1^{st}$, $5^{th}$, $20^{th}$ and $40^{th}$ analogs for the Ulrichen station on the whole calibration period, due to introduction of the MTW. The precipitation target remains as before, that is centered on 18~h~UTC (6~h~UTC to 6~h~UTC the next day). The shapes of the distributions of the conventional approach and the MTW are similar, but the values of the analogy criteria are now reduced (shifted to the left), and therefore better. An increase in the difference between a fixed window and a moving window is identifiable, which means that the last analogs are further improved. The latter effect is due to the accumulation of improvements brought by the new analog situations in the selection.

The improvements of the $S_{\text{TW}}$ criteria are summarized in Fig. \ref{fig:changes_S1}, which shows (top) quantiles of the $S_{\text{TW}}$ criteria according to the analog rank for the conventional method and the MTW, and (bottom) quantiles of the relative reduction. This confirms that all quantile seem similarly reduced ($S_{\text{TW}}$ distributions keep their shape), and that this improvement is constantly increasing from the first to the last analog (Fig. \ref{fig:changes_S1} bottom).

The median of the $S_{\text{TW}}$ values reduction (Fig. \ref{fig:changes_S1} bottom) starts approximately at 5~\% for the first analog and reaches more than 10~\% for the last one. This increasing trend with the analog rank can be explained by the accumulation of better analogs in the distribution. The minimum improvement starts from 0 and reaches 2--3\%, meaning that the criteria have been improved on most analogs for every day of the calibration period. All other stations have a similar improvement of the $S_{\text{TW}}$ criteria, both in terms of distribution shape and amplitude.

@@

A seasonal effect can be observed on the reduction of the $S_{\text{TW}}$ criteria due to the MTW (Fig. \ref{fig:changes_S1_seasons}). The improvements are greater for winter than summer. One hypothesis is that the diurnal effects of the summer months have an influence on the atmospheric circulation, at least in the lower layers. This effect is based on the daily cycle and good analogs are essentially found for the same hours or time windows.

An analysis of the selected hours for the geopotential predictor seems to confirm this assumption (Fig. \ref{fig:hours_selection_per_season}). It was found that the new choice of the temporal window in winter, when using the MTW approach, is well balanced between the 4 options. This means a change of 75\% of the analogs selection compared to the conventional approach, which improves the circulation analogy.

On the contrary, the summer months have a preference for the initial temporal window (Z500 24~h \& Z1000 12~h), due to more pronounced diurnal effects which reduces the potential for improvement of the criteria. Other seasons are between these two extremes, which is consistent with their respective improvements. This seasonal effect was observed for each station in a very similar way, and generally even with a slightly larger amplitude than for Ulrichen (see example of the Visp station in Fig. \ref{fig:hours_selection_per_season_Visp}).

@@

When considering the second level of analogy of the 2Z-2MI method (Table \ref{table:method_2Z-2MI}), the number of candidates situations did not increase, as they are subsampled in the previously $N_{1}$ selected analogs, but their dates have changed. In contrast to earlier, both a reduction or an increase of the $E_{\text{RMS}}$ analogy criterion values are possible. Indeed, Fig. \ref{fig:changes_RMSE} shows an almost insignificant improvement of the $E_{\text{RMS}}$ values. Unlike for the first level of analogy, the relative improvement of the $E_{\text{RMS}}$ values are distributed relatively symmetrically around zero, with improvements and losses of the same amplitude. Once again, the results for the other stations are similar.

This result of a globally null improvement of the $E_{\text{RMS}}$ values does not mean that the 2Z-2MI method cannot be improved by the MTW. It means that after the selection of the analogs situations in terms of the synoptic circulation, the new candidate dates do not allow to find better analogs in terms of moisture. However, the selected dates have changed in the first level of analogy, and thus also in the final selection, which can potentially improve the performance scores on the predictand.

@@

A systematic improvement of the $S_{\text{TW}}$ values was previously observed. However, finding better analog situations does not obligatory imply better skills to predict the precipitation. Therefore, the impact of the MTW introduction, and thus the selection of other analog dates, has to be assessed on the performance scores. In order to perform this assessment, the 24~h-totals (moving average) at a 6-hourly time step were used (see Sect. \ref{sec:data}). The target dates remain unchanged, since the original time slot (6~h~UTC -- 6~h~UTC the next day) is kept, and thus the performance scores can be directly compared with the former ones.

@@

After recalibrating, some changes in the optimal parameters can be observed for both 2Z (Table \ref{table:params_2Z_new}) and 2Z-2MI methods (Table \ref{table:params_2Z-2MI_new}). Among these, the east-west dimension of the spatial windows of the circulation analogy tends to decrease. More significantly, the optimal number of analogs increases after introducing the MTW, of a significant order of magnitude: 25~\% to 83~\% for the 2Z method and 20~\% 67~\% for the 2Z-2MI method. The number of analogs of the first analogy level of the 2Z-2MI method even reached three times its previous value for the Visp station. It seems as if the method perceive the increase in the archive length and diversity by allowing to extract more analogs. As we saw in Fig. \ref{fig:changes_S1}, the improvement of the $S_{\text{TW}}$ criteria grows along with the rank of the analog, which shows an accumulation of better analog situations in the distributions. It seems thus profitable to widen the selection of analogs in order to keep also some whose rank has increased, as they appear to be also relevant to predict precipitation values. The number of good analogs is thus globally increased.

This increase in number of analogs has a slight effect on the performance of the different precipitation classes. The same analysis as in Sect. \ref{sec:improvement_CRPSS_precip_threshold} has been performed again on the basis of the newly calibrated parameters. The results are generally very similar, but a slight performance increase of small precipitation values can be observed at the expense of higher amounts. The change in the analogs numbers is likely to be responsible for this difference in the performance distributions.

@@

A proportional distribution is certainly the simplest approach that can be performed. It consists in allocating proportional parts of the original daily time series into a new moving average of 24~h-totals (Fig. \ref{fig:illustration_disaggregation}). When using this reconstructed time series on the shorter period, the method performance was degraded (Table \ref{table:disaggregation_proportional} to compare to Table \ref{table:CRPSS_MTW}) and was even below the conventional method without MTW (Table \ref{table:loss_reduction}). The benefit of a better selection of the analog situations is lost due to a precipitation archive of poor quality.

@@

The first step consisted in determining which variables from the precipitable water or the relative humidity (at 1000~hPa, 925~hPa or 850~hPa) produces the precipitation chronology that is the most correlated (on non-zero values) with the accurate 6-hourly precipitation time series on the period 1982-2007. The location of the most predictive point from the reanalysis grid had to be assessed (closest points: 5--7.5° E and 45--47.5° N), as well an eventual time lapse between both series, due to the significant distance separating the weather stations and the reanalysis grid point. The most relevant point will probably not be the same for all stations, since one can expect to find the optimum in the direction of the main moisture inflows for the station. This assessment must be carried out for each station.

The results of the correlation analyses (on non-zero values) between the new series reconstituted using the proxy and the actual series are illustrated in Table \ref{table:proxy_correlations} for Zermatt (on the short period, 1982--2007). The best proxy is the precipitable water at 45° N and 7.5° E, with a time offset of -6~h. The determination coefficient of 0.721 is higher than that of the series from the proportional distribution (0.698), which confirms that a little information was added to the series of precipitation.

@@

The AM is most often based on a daily time step, due to the availability of long precipitation archives. However, it is unlikely that two analog synoptic situations, that evolve relatively quickly, would correspond optimally at the same hour of the day. It is thus more likely that a better match can be found at another time, which can change the selection of the analogs dates.

As \citet{Finet2008} had previously shown, the introduction of a MTW allows finding better analog situations in terms of the atmospheric circulation. It has been here demonstrated that the improvement of the $S_{\text{TW}}$ criteria values is growing with the rank of the analog. This is due to the accumulation of better analogs within the predicted distributions.

@@

A seasonal effect has been highlighted, as the MTW is more profitable for winter months than summer. The reason is likely that the diurnal cycle has a bigger effect in summer than in winter, which results in better analogs at the same time of the day. The preference for the same hours in summer has been demonstrated, whereas 75\% of the analog situations are selected at a different time in winter.

@@

The parameters were then calibrated again, using the MTW. Some parameters changed, the main difference being the number of analogs, which systematically and significantly increased compared to the original set. A wider selection of analog situations, containing those whose rank decreased, seemed profitable for the prediction performance. The number of good analogs is thus globally increased. This change seems to benefit mainly the prediction of days with small precipitation totals.

The importance of the quality of the precipitation archive was also demonstrated, as too simplistic reconstructions of 6-hourly time series lead to a loss of all the improvement brought by better analog situations identified by the MTW. The precipitation prediction is improved only when the precipitation chronology is close to the accurate one. Attempts to reconstruct longer time series based on a simplistic proportional distribution or using meteorological variables from the NCEP reanalysis 1 dataset as proxy did not succeed. Other reanalysis datasets could eventually preform better.

@@

\begin{figure*}[htb]
	\begin{center}
		\includegraphics[width=15cm]{figures/changes_S1_analogues.pdf}
	\end{center}
	\caption{Changes in the $S_{\text{TW}}$ criteria distributions of (a) the $1^{st}$, (b) $5^{th}$, (c) $20^{th}$ and (d) $40^{th}$ analog ranks for the Ulrichen station, due to the MTW.}
	\label{fig:changes_S1_analogs}
\end{figure*}

\begin{figure}[htb]
	\begin{center}
		\includegraphics[width=8.2cm]{figures/changes_S1_value_and_gain.pdf}
	\end{center}
	\caption{Synthesis of the changes in the $S_{\text{TW}}$ criteria, due to the MTW, for the Ulrichen station, depending on the ranks of the analog. (a) Quantiles of the $S_{\text{TW}}$ distributions with and without the MTW. (b) Quantiles of the relative improvements of the S1 criteria when using the MTW.}
	\label{fig:changes_S1}
\end{figure}

@@

\begin{figure}[htb]
	\includegraphics[width=8.3cm]{figures/hours_selection_per_season.pdf}
	\caption{Distribution of the predictors hours in the selected analog dates depending on the season, for the Ulrichen station.}
	\label{fig:hours_selection_per_season}
\end{figure}

\begin{figure}[htb]
	\includegraphics[width=8.3cm]{figures/hours_selection_per_season_Visp.pdf}
	\caption{Distribution of the predictors hours in the selected analog dates depending on the season, for the Visp station.}
	\label{fig:hours_selection_per_season_Visp}
\end{figure}

\begin{figure}[htb]
	\begin{center}
		\includegraphics[width=8.1cm]{figures/changes_RMSE_value_and_gain.pdf}
	\end{center}
	\caption{Synthesis of the changes in the $E_{\text{RMS}}$ criteria, due to the MTW, for the Ulrichen station, depending on the ranks of the analog. (a) Quantiles of the $E_{\text{RMS}}$ distributions with and without the MTW. (b) Quantiles of the relative improvements of the $E_{\text{RMS}}$ criteria when using the MTW.}
	\label{fig:changes_RMSE}
\end{figure}
